Fig3

### 2020-05-25 analysis/06_datawrangling.Rmd
new genes lists, but candidate vsd is empty :(

```{r setup}
library(tidyverse)
library(readr)
library(corrr)

source("../R/themes.R")
source("../R/functions.R")
```

@@

```{r allvsd, message = F, warning  = F}
vsd_path <- "../results/DEseq2/hypothesis/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

@@

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = blk.s061.pu.y_female_gonad_inc.d9:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts") 
allvsd %>% select(-file_name) %>% head()

head(allvsd)
```

@@

# candidate genes from GO and literature
parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% pull(gene)
parentalcaregenes

## genes WGCNA prl module
WGCNAgenes <- read_csv("../results/05_PRLmodule.csv") %>% pull(x)
WGCNAgenes

unique(allDEG$comparison)

## hilogenes

@@

eggchickgenes <- allDEG %>%
  filter(comparison != "lo_hi")  %>%
  drop_na() %>%
  arrange(padj)  %>% 
  top_n(20) %>% pull(gene)

@@

```{r candidatevsd}
candidategenes <- c(parentalcaregenes, WGCNAgenes, hilogenes, eggchickgenes)

getcandidatevsd2 <- function(whichgenes, whichtissue, whichsex){
  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/hypothesis/'), "[", 2)) %>%
    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                  tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                  treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
    filter(tissue == whichtissue, sex %in% whichsex)  %>%
    drop_na()
  #candidates$treatment <- factor(candidates$treatment, levels = alllevels)
  return(candidates)
}

hypvsd <- getcandidatevsd2(candidategenes, "hypothalamus", sexlevels)
pitvsd <- getcandidatevsd2(candidategenes, "pituitary", sexlevels)
gonvsd <- getcandidatevsd2(candidategenes, "gonad", sexlevels)
candidatevsd <- rbind(hypvsd, pitvsd)
candidatevsd <- rbind(candidatevsd, gonvsd)

head(candidatevsd)
```

```{r write files}
write.csv(candidatevsd, "../results/06_candidatevsd.csv", row.names = F)
write.csv(allDEG, "../results/06_allDEG.csv", row.names = F)

>>>>

```{r setup}
library(tidyverse)
library(readr)
library(corrr)
library(forcats)

source("../R/themes.R")
source("../R/functions.R")
source("../R/genelists.R")
```

@@

```{r allvsd, message = F, warning  = F}
vsd_path <- "../results/DEseq2/treatment/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

@@

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts")

allvsd %>%  select(-file_name) %>% head()

colData %>% filter(tissue == "gonads", treatment == "bldg")

allvsd %>% filter(samples == "blk11.x_female_gonad_bldg")
```

@@

## "internal clock" PRL genes

@@

## external environment gens 
eggchickgenes <- allDEG %>%
  filter(comparison != "lo_hi")  %>%
  drop_na() %>%
  arrange(padj)  %>% 
  top_n(20) %>% pull(gene)

@@

```{r candidatevsd}
# all list of candidate gens accumlated here

candidategenes <- c(parentalcaregenes, WGCNAgenes, hilogenes, eggchickgenes, 
                    suszynskaagenes, shaidgenes)

getcandidatevsd2 <- function(whichgenes, whichtissue, whichsex){
  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/treatment/'), "[", 2)) %>%
    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                  tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                  treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
    filter(tissue == whichtissue, sex %in% whichsex)  %>%
    drop_na()
  #candidates$treatment <- factor(candidates$treatment, levels = alllevels)
  return(candidates)
}

hypvsd <- getcandidatevsd2(candidategenes, "hypothalamus", sexlevels)
pitvsd <- getcandidatevsd2(candidategenes, "pituitary", sexlevels)
gonvsd <- getcandidatevsd2(candidategenes, "gonad", sexlevels)

candidatevsd <- rbind(hypvsd, pitvsd)
candidatevsd <- rbind(candidatevsd, gonvsd) 
candidatevsd <- candidatevsd %>%
  mutate(treatment = factor(treatment)) %>%
  mutate(treatment = fct_recode(treatment, "early" = "m.inc.d8" )) %>%
  mutate(treatment = factor(treatment, alllevels))
tail(candidatevsd)
```

```{r writefiles}
write.csv(candidatevsd, "../results/06_candidatevsd.csv", row.names = F)
write.csv(allDEG, "../results/06_allDEG.csv", row.names = F)
```

### 2020-05-25 analysis/figure-3.Rmd
new genes lists, but candidate vsd is empty :(

source("../R/themes.R") 
source("../R/functions.R")

@@

pca1f <- subsetmakepca("pituitary", charlevelsnocontrol, "female")	
pca2f <- makefvizdf("pituitary", charlevelsnocontrol, "female")

pca1m <- subsetmakepca("pituitary", charlevelsnocontrol, "male")	
pca2m <- makefvizdf("pituitary", charlevelsnocontrol, "male")

pca1 <- subsetmakepca("pituitary", charlevelsnocontrol, sexlevels)	
pca2 <- makefvizdf("pituitary", charlevelsnocontrol, sexlevels)	
```

@@

```{r}
# candidate counts for boxplots
candidatevsd <- read_csv("../results/06_candidatevsd.csv") %>%
  mutate(treatment = factor(treatment)) %>%
  mutate(tissue = factor(tissue, levels = tissuelevel),
        treatment = factor(treatment, levels = charlevelsnocontrol)) %>%
  drop_na()

@@

# DEGs infor for PRL and external
hilodf <- allDEG %>%
  filter(comparison == "lo_hi")  %>%
  arrange(padj)

eggchickdf <- allDEG %>%
  filter(!comparison %in% c("lo_hi", "nest_eggs", "nest_chicks"))  %>%
  arrange(padj)

eggchickgenes <- eggchickdf %>%  top_n(20) %>% pull(gene)
hilogenes <- hilodf %>%  top_n(20) %>% pull(gene)

## for correlations
pithilo <- makecorrdf("female", "pituitary", hilogenes)
pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[1:20])

## for scatter correlation

candidatevsdwide <- candidatevsd  %>%
    pivot_wider(names_from = gene, values_from = counts) 
pitwide <- subsetcandidatevsdwide(sexlevels, "pituitary")

@@

c1 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 lo vs. hi PRL DEGs") 
c2 <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes from WGCNA module with PRL") + theme(legend.position = "right")
c3 <- candidateboxplot("pituitary", c("KPNA2"), sexlevels)
c4 <- candidateboxplot("pituitary", c("CDK1"), sexlevels) + 
  theme(axis.text.x = element_text(angle = 45, hjust =1), strip.text = element_blank())  + labs(x = " ")
c34 <- plot_grid(c3,c4, nrow = 2, rel_heights = c(1,1.2))

c <- plot_grid(c1,c2,c34, nrow = 1, labels = c("D", "", "E"), label_size = 8, rel_widths = c(1.1,1.4,1))

@@

pdf(file="../figures/fig3-1.pdf", width=7.25, height=7.25)
plot(fig3)
dev.off()
```

```{r savefiles}
write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)

@@

>>>>

source("../R/themes.R") 
source("../R/functions.R")
source("../R/genelists.R")

@@

pca1f <- subsetmakepca("pituitary", charlevels, "female")	
pca2f <- makefvizdf("pituitary", charlevels, "female")

pca1m <- subsetmakepca("pituitary", charlevels, "male")	
pca2m <- makefvizdf("pituitary", charlevels, "male")

pca1 <- subsetmakepca("pituitary", charlevels, sexlevels)	
pca2 <- makefvizdf("pituitary", charlevels, sexlevels)	
```

@@

```{r}
# candidate counts for boxplots
candidatevsd <- read_csv("../results/06_candidatevsd.csv") %>%
  mutate(treatment = factor(treatment)) %>%
  mutate(tissue = factor(tissue, levels = charlevels),
        treatment = factor(treatment, levels = charlevels)) %>%
  drop_na()
head(candidatevsd)

@@

# DEGs infor for PRL and external
hilodf <- allDEG %>%
  filter(comparison == "lo_hi")  %>%
  arrange(padj) 
hilogenes <- hilodf %>%  top_n(20) %>% pull(gene)

eggchickdf <- allDEG %>%
  filter(!comparison %in% c("lo_hi", "nest_eggs", "nest_chicks"))  %>%
  arrange(padj)  
eggchickgenes <- eggchickdf %>%  top_n(20) %>% pull(gene)

## data-driven PRL genes
pithilo <- makecorrdf("female", "pituitary", hilogenes)
pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[1:20])

## https://www.gynecologiconcology-online.net/article/S0090-8258(19)30069-1/fulltext
cancergenes <- makecorrdf("female", "pituitary", c(shaidgenes,suszynskaagenes))

## for scatter correlation (not currently usesed)
candidatevsdwide <- candidatevsd  %>%
    pivot_wider(names_from = gene, values_from = counts) 
pitwide <- subsetcandidatevsdwide(sexlevels, "pituitary")

@@

c1 <- plotcorrplot(pithilo, NULL) + theme(legend.position = "none") + labs(subtitle = "Top 20 lo vs. hi PRL DEGs") 
c2 <- plotcorrplot(pitWGCNA, NULL) + theme(legend.position = "none") + labs(subtitle = "20 genes from WGCNA module with PRL") 
c3 <- plotcorrplot(cancergenes, NULL) + theme(legend.position = "right") + labs(subtitle = "Breast, ovary, piuitary cancer genes") 
c <- plot_grid(c1,c2,c3, nrow = 1, labels = c("D", "E", "F"), label_size = 8, rel_widths = c(1,1,1.4))
c

@@

@@

```{r savefiles}
#write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)

#pdf(file="../figures/fig3-1.pdf", width=7.25, height=7.25)
#plot(fig3)
#dev.off()
```

```{r forvictoria, eval = F}
## PRL and PRLR in three tissues

candidatevsd <- candidatevsd %>%
  filter(treatment %in%  c("early", "extend", charlevels))

p1 <- candidateboxplot("hypothalamus", c("PRL"), sexlevels) + labs(subtitle = "hypothalamus", title = "PRL")
p2 <- candidateboxplot("pituitary", c("PRL"), sexlevels) + labs(subtitle = "pituitary")
p3 <- candidateboxplot("gonad", c("PRL"), sexlevels) + labs(subtitle = "gonads") + theme(axis.text.x = element_text(angle = 45))

p4 <- candidateboxplot("hypothalamus", c("PRLR"), sexlevels) + labs(subtitle = " ", title = " ") 
p5 <- candidateboxplot("pituitary", c("PRLR"), sexlevels) + labs(subtitle = " ")
p6 <- candidateboxplot("gonad", c("PRLR"), sexlevels) + theme(axis.text.x = element_text(angle = 45))+ labs(subtitle = " ")

plot_grid(p1,p4,p2,p5,p3,p6, ncol = 2, rel_heights = c(1.2,1,1.2))
```
